---
title: "tudy"
author: "JiyueQin"
date: "November 19, 2018"
output: github_document
---


Risk factors that we're interested in:: (15 total)

* ACCESS2: Current lack of health insurance among adults aged 18<89><db><d2>64 Years
* BINGE: Binge drinking among adults aged >=18 Year
* BPHIGH: High blood pressure among adults aged >=18 Year
* BPMED: Taking medicine for high blood pressure control among adults aged >=18 Years with high blood pressure
* OBESITY: Obesity among adults aged >=18 Years
* CHECKUP: Visits to doctor for routine checkup within the past Year among adults aged >=18 Years
* CHOLSCREEN: Cholesterol screening among adults aged >=18 Years
* CSMOKING: Current smoking among adults aged >=18 Years
* DIABETES: Diagnosed diabetes among adults aged >=18 Years
* HIGHCHOL: High cholesterol among adults aged >=18 Years who have been screened in the past 5 Years
* KIDNEY: Chronic kidney disease among adults aged >=18 Years
* LPA: No leisure-time physical activity among adults aged >=18 Years
* MHLTH: Mental health not good for >=14 days among adults aged >=18 Years
* PHLTH: Physical health not good for >=14 days among adults aged >=18 Years
* SLEEP: Sleeping less than 7 hours among adults aged >=18 Years



Outcomes that we're interested in:: (2 total)

* CHD: Coronary heart disease among adults aged >=18 Years
* STROKE: Stroke among adults aged >=18 Years




```{r, include=FALSE}
  library(tidyverse)
  library(stringr)
  library(tidyverse)
  library(purrr)
  library(viridis)
  library(ggrepel)
  ##library(ggmap)
  library(plotly)
```

## load the data, not evaluating this
```{r cache = TRUE, eval = FALSE}
cvrisk = read_csv("https://data.cdc.gov/api/views/6vp6-wxuq/rows.csv?accessType=DOWNLOAD") %>% 
    janitor::clean_names() 


## data used for shiny
##cvrisk = read_csv("https://chronicdata.cdc.gov/api/views/6vp6-wxuq/rows.csv?accessType=DOWNLOAD")

```

## load data 
```{r cache = TRUE}
cvrisk_url = "https://data.cdc.gov/api/views/6vp6-wxuq/rows.csv?accessType=DOWNLOAD"

## Read in data from github
cvrisk = 
  read.csv(url(cvrisk_url)) %>% 
  janitor::clean_names() %>% 
  as_tibble()

```




```{r}
## Annie EDA - final set of tidying data

## create lookup table for measures and their meaning

cvrisk_measure_lookup = 
  cvrisk %>% 
  group_by(measure, measure_id, short_question_text, year) %>% 
  count() %>% 
  arrange(year, measure_id)


## clean up nyc datset (only include 2015 data, and remove missing values)

nyc_cvrisk = 
  cvrisk %>% 
  filter(state_desc == "New York", city_name == "New York",
         geographic_level == "Census Tract",
         !is.na(data_value),
         ##year == 2016,
         measure_id %in% c("ACCESS2", "BINGE", "BPHIGH", "BPMED", "OBESITY", "CHECKUP", 
                         "CHOLSCREEN", "CSMOKING", "DIABETES", "HIGHCHOL", "KIDNEY", 
                         "LPA", "MHLTH", "PHLTH", "SLEEP", "CHD", "STROKE")) %>% 
  separate(geo_location, into = c("latitude", "longitude"), sep = ",") %>% 
  mutate(latitude = str_replace(latitude, "\\(", ""),
         longitude = str_replace(longitude, "\\)", "")) %>% 
  select(unique_id, latitude, longitude, population_count, measure_id, data_value, short_question_text)


nyc_cvrisk_wide = 
  nyc_cvrisk %>% 
  spread(key = measure_id, value = data_value) %>% 
  janitor::clean_names()


nyc_cvrisk %>%
  ggplot(aes(x = measure_id, y = data_value)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


Exploratory -- plotting each outcome by each risk factor
```{r}

## nyc plots, each risk factor x chd
nyc_cvrisk_wide %>% 
  ggplot(aes(x = access2, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = binge, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = bphigh, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = bpmed, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = checkup, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = cholscreen, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = csmoking, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = diabetes, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = highchol, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = kidney, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = lpa, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = mhlth, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = obesity, y = chd)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = phlth, y = chd)) + 
  geom_point()



## nyc plots, each risk factor x stroke


nyc_cvrisk_wide %>% 
  ggplot(aes(x = access2, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = binge, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = bphigh, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = bpmed, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = checkup, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = cholscreen, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = csmoking, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = diabetes, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = highchol, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = kidney, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = lpa, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = mhlth, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = obesity, y = stroke)) + 
  geom_point()

nyc_cvrisk_wide %>% 
  ggplot(aes(x = phlth, y = stroke)) + 
  geom_point()


```

Correlation between risk factors/outcomes
```{r}
nyc_cvrisk_wide %>% 
  select(access2:stroke) %>% 
  cor()

## install.packages("corrplot")
library(corrplot)

nyc_cvrisk_wide %>% 
  select(access2:stroke) %>% 
  cor()%>%
  corrplot(., method="circle")

```

```{r}
library(patchwork)
library(tidyverse)


access2_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = access2, y = chd)) + 
  geom_point()

binge_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = binge, y = chd)) + 
  geom_point()

bphigh_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = bphigh, y = chd)) + 
  geom_point()

bpmed_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = bpmed, y = chd)) + 
  geom_point()

checkup_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = checkup, y = chd)) + 
  geom_point()

cholscreen_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = cholscreen, y = chd)) + 
  geom_point()

csmoking_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = csmoking, y = chd)) + 
  geom_point()

diabetes_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = diabetes, y = chd)) + 
  geom_point()

highchol_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = highchol, y = chd)) + 
  geom_point()

kidney_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = kidney, y = chd)) + 
  geom_point()

lpa_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = lpa, y = chd)) + 
  geom_point()

mhlth_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = mhlth, y = chd)) + 
  geom_point()

obesity_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = obesity, y = chd)) + 
  geom_point()

phlth_plot = nyc_cvrisk_wide %>% 
  ggplot(aes(x = phlth, y = chd)) + 
  geom_point()


##internal
(phlth_plot + obesity_plot + mhlth_plot + kidney_plot) / 
  (highchol_plot + diabetes_plot+ bphigh_plot)


(phlth_plot + obesity_plot) / 
  (highchol_plot + diabetes_plot)

##external
(lpa_plot + csmoking_plot + cholscreen_plot + checkup_plot) / 
  (bpmed_plot + binge_plot + access2_plot)

##total
(phlth_plot + obesity_plot + mhlth_plot + lpa_plot) / 
  (kidney_plot + highchol_plot + diabetes_plot  + csmoking_plot) / 
  (cholscreen_plot + checkup_plot + bpmed_plot + bphigh_plot) / 
  (binge_plot + access2_plot)


```



```{r}
## gganimate packages and day tidying::::


## this is the old package - can use gganimate with this package
devtools::install_github("nteetor/gganimate")

## new gganimate package
install.packages("cowplot")
devtools::install_github("dgrtwo/gganimate")

## dataset used in examples
install.packages("gapminder")

library("dgrtwo/gganimate")
library(tidyverse)
library(ggplot2)
library(cowplot)
library(gganimate)
##library(installr)

## have not gotten ImageMagick to work correctly
install.packages("installr")
install.ImageMagick()

## create table for animation
chd_by_id = nyc_cvrisk %>% 
  group_by(unique_id) %>% 
  summarise(chd_value = max(data_value)) 

nyc_cvrisk_joined = left_join(nyc_cvrisk, chd_by_id, by = "unique_id")



```


```{r}
## gganimate attempts


theme_set(theme_bw())  # pre-set the bw theme.


## using transition_length and state_length
p <- 
  ggplot(data = nyc_cvrisk_joined, aes(x = chd_value, y = data_value, frame = short_question_text)) +
  geom_point(data = nyc_cvrisk_joined, aes( 
                                           size = population_count, colour = short_question_text ),
             alpha = 0.5) + 
  ggtitle(nyc_cvrisk_joined$short_question_text) +
  
  labs(
       x = 'Coronary Heart Prevalence', 
       y = 'Risk Factor Prevalence') +
  
  # gganimate parts
  transition_states(short_question_text, transition_length = 1, state_length = 50000, wrap = TRUE) +
  enter_fade() +
  exit_fade()

animate(p, fps = 1)


## title = 'Risk Factor: {measure_id}', 

## add title
## facet by stroke (remove stroke and chd)
## add title
## slow down
## make points more transparent
## change order




