---
title: "Exploration - MG"
author: "Margaret Gacheru"
date: "November 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(leaps)
library(lattice)
library(corrplot)

```


Load the data 
```{r cache = TRUE}
cv_data = read_csv("https://data.cdc.gov/api/views/6vp6-wxuq/rows.csv?accessType=DOWNLOAD")
```


Limiting database to New York 

```{r, warning=FALSE, message=FALSE}
nyc_cvrisk = cv_data %>% 
  janitor::clean_names()%>%
  filter(state_desc != "United States", geographic_level != "City")%>%
  filter(year == 2015, state_desc == "New York")%>%
  select(unique_id, state_desc, city_name, population_count, geo_location, short_question_text, data_value)%>%
  spread(key = short_question_text, value = data_value)%>%
  separate(geo_location, into = c("latitude", "longitude"), sep = ",")%>%
  mutate(latitude = str_replace(latitude, "\\(", ""),
         latitude = str_replace(latitude, " ", ""),
         latitude = as.numeric(latitude),
         longitude = str_replace(longitude, "\\)", ""),
         longitude = str_replace(longitude, " ", ""),
         longitude = as.numeric(longitude))%>%
  janitor::clean_names()
  
```


Avoidable behaviours (physical inactivity, current smoking) in New York

```{r}

nyc_cvrisk%>%
  filter(city_name == "New York")%>%
  ggplot(aes(x = physical_inactivity, y = current_smoking, color=coronary_heart_disease)) + 
  geom_point(aes(size = coronary_heart_disease), alpha = .4) +
  labs(
    y = "Physical Activity",
    x = "Current Smoking"
  )+
  viridis::scale_color_viridis(discrete = FALSE)

```


Top 10 cities with coronary heart disease (external factors)

```{r}

top = nyc_cvrisk%>%
  group_by( city_name)%>%
  summarize(Outcome = median(coronary_heart_disease),
            Checkup = median(annual_checkup),
            Cholesterol_Screen = median(cholesterol_screening),
            Smoking = median(current_smoking),
            Insurance = median(health_insurance),
            Inactivity = median(physical_inactivity),
            Binge_Drinking = median(binge_drinking))%>%
  arrange(desc(Outcome))%>%
  head(10)

row.names(top) = top$city_name

top_matrix = top%>%select(-city_name)%>%data.matrix(top) 

levelplot(top_matrix, col.regions=heat.colors, colorkey=list(space="right"),
          xlab = "", ylab = "", main = "Top 10 Cities with Coronary Heart Disease (External Factors)")
```


Bottom 10 cities with coronary heart disease (external factors)

```{r}

bottom = nyc_cvrisk%>%
  group_by( city_name)%>%
  summarize(Outcome = median(coronary_heart_disease),
            Checkup = median(annual_checkup),
            Cholesterol_Screen = median(cholesterol_screening),
            Smoking = median(current_smoking),
            Insurance = median(health_insurance),
            Inactivity = median(physical_inactivity),
            Binge_Drinking = median(binge_drinking))%>%
  filter(Outcome != "NA")%>%
  arrange(Outcome)%>%
  head(10)

row.names(bottom) = bottom$city_name

bottom_matrix = bottom%>%select(-city_name)%>%data.matrix(bottom) 

library(lattice)
levelplot(bottom_matrix, col.regions=heat.colors, colorkey=list(space="right"),
          xlab = "", ylab = "", main = "Bottom 10 Cities with Coronary Heart Disease (External Factors)")

```


Top 10 cities with coronary heart disease (internal factors)

```{r}

top = nyc_cvrisk%>%
  group_by( city_name)%>%
  summarize(Outcome = median(coronary_heart_disease),
            Arthritis = median(arthritis),
            Kidney_Disease = median(chronic_kidney_disease),
            COPD = median(copd),
            Diabetes = median(diabetes),
            High_Blood_Pressure = median(high_blood_pressure),
            High_Cholesterol = median(high_cholesterol),
            Stroke = median(stroke))%>%
  arrange(desc(Outcome))%>%
  head(10)

row.names(top) = top$city_name

top_matrix = top%>%select(-city_name)%>%data.matrix(top) 

levelplot(top_matrix, col.regions=heat.colors, colorkey=list(space="right"),
          xlab = "", ylab = "", main = "Top 10 Cities with Coronary Heart Disease (Internal Factors)")
```


Bottom 10 cities with coronary heart disease (internal factors)

```{r}

bottom = nyc_cvrisk%>%
  group_by( city_name)%>%
  summarize(Outcome = median(coronary_heart_disease),
            Arthritis = median(arthritis),
            Kidney_Disease = median(chronic_kidney_disease),
            COPD = median(copd),
            Diabetes = median(diabetes),
            High_Blood_Pressure = median(high_blood_pressure),
            High_Cholesterol = median(high_cholesterol),
            Stroke = median(stroke))%>%
  filter(Outcome != "NA")%>%
  arrange(Outcome)%>%
  head(10)

row.names(bottom) = bottom$city_name

bottom_matrix = bottom%>%select(-city_name)%>%data.matrix(bottom) 

library(lattice)
levelplot(bottom_matrix, col.regions=heat.colors, colorkey=list(space="right"),
          xlab = "", ylab = "", main = "Bottom 10 Cities with Coronary Heart Disease (Internal Factors)")

```

It is difficult to visually see the relationships between the variables (even when there is a high correlation) because the range in each variable is fairly small


Correlation matrix (for covariates and outcome)

```{r}

library(corrplot)
res = nyc_cvrisk%>%
  select(-c("unique_id", "state_desc","city_name", "population_count", "latitude", "longitude"))%>%
  drop_na()%>%
  cor()

corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 90)

ggplot(nyc_cvrisk, aes(x = health_insurance, y = coronary_heart_disease))+
  geom_point()
```


Cleaner correlation matrix (CHD)

```{r}
cv_data = read_csv("https://data.cdc.gov/api/views/6vp6-wxuq/rows.csv?accessType=DOWNLOAD")


## clean up nyc datset (only include 2016 data, and remove missing values)

ny_cvrisk_2016 = 
  cv_data %>% 
  janitor::clean_names()%>%
  filter(state_desc == "New York",
         geographic_level == "Census Tract",
         !is.na(data_value),
         year == 2016,
         measure_id %in% c("ACCESS2", "BINGE", "BPHIGH", "BPMED", "OBESITY", "CHECKUP", 
                         "CHOLSCREEN", "CSMOKING", "DIABETES", "HIGHCHOL", "KIDNEY", 
                         "LPA", "MHLTH", "PHLTH", "SLEEP", "CHD")) %>% 
  separate(geo_location, into = c("latitude", "longitude"), sep = ",") %>% 
  mutate(latitude = str_replace(latitude, "\\(", ""),
         longitude = str_replace(longitude, "\\)", "")) %>% 
  select(unique_id, latitude, longitude, population_count, measure_id, data_value)


ny_cvrisk_2016_wide = 
  ny_cvrisk_2016 %>% 
  spread(key = measure_id, value = data_value) %>% 
  janitor::clean_names()


ny_cvrisk_2016_wide %>% 
  select(access2:sleep) %>% 
  select("chd", everything())%>%
  cor() %>% 
  corrplot(., method="circle")



```


Cleaner correlation matrix (Stroke)

```{r}

## clean up nyc datset (only include 2016 data, and remove missing values)

ny_cvrisk_2016 = 
  cv_data %>% 
  janitor::clean_names()%>%
  filter(state_desc == "New York",
         geographic_level == "Census Tract",
         !is.na(data_value),
         year == 2016,
         measure_id %in% c("ACCESS2", "BINGE", "BPHIGH", "BPMED", "OBESITY", "CHECKUP", 
                         "CHOLSCREEN", "CSMOKING", "DIABETES", "HIGHCHOL", "KIDNEY", 
                         "LPA", "MHLTH", "PHLTH", "SLEEP", "STROKE")) %>% 
  separate(geo_location, into = c("latitude", "longitude"), sep = ",") %>% 
  mutate(latitude = str_replace(latitude, "\\(", ""),
         longitude = str_replace(longitude, "\\)", "")) %>% 
  select(unique_id, latitude, longitude, population_count, measure_id, data_value)


ny_cvrisk_2016_wide = 
  ny_cvrisk_2016 %>% 
  spread(key = measure_id, value = data_value) %>% 
  janitor::clean_names()


corr_matrix = ny_cvrisk_2016_wide %>% 
  select(access2:stroke) %>% 
  select("stroke", everything())%>%
  cor()%>%View()

colnames(corr_matrix) <- c("Stroke", "No Health Insurance", "Binge Drinking", "Routine Checkup",
                          "Smoking", "Diabetes", "Kidney Disease", "No Physical Activity", "Poor Mental Health", "Obesity", "Poor Physical Health", "Poor Sleep")

corrplot(corr_matrix, method="circle")


M <- cor(mtcars)[1:5,1:5]%>%View()
colnames(M) <- c("alpha", "beta", ":alpha+beta", ":a[0]", "=a[beta]")
rownames(M) <- c("alpha", "beta", NA, "$a[0]", "$ a[beta]")
corrplot(M)

```

* ACCESS2: Current lack of health insurance among adults aged 18<89><db><d2>64 Years
* BINGE: Binge drinking among adults aged >=18 Year
* BPHIGH: High blood pressure among adults aged >=18 Year
* BPMED: Taking medicine for high blood pressure control among adults aged >=18 Years with high blood pressure
* OBESITY: Obesity among adults aged >=18 Years
* CHECKUP: Visits to doctor for routine checkup within the past Year among adults aged >=18 Years
* CHOLSCREEN: Cholesterol screening among adults aged >=18 Years
* CSMOKING: Current smoking among adults aged >=18 Years
* DIABETES: Diagnosed diabetes among adults aged >=18 Years
* HIGHCHOL: High cholesterol among adults aged >=18 Years who have been screened in the past 5 Years
* KIDNEY: Chronic kidney disease among adults aged >=18 Years
* LPA: No leisure-time physical activity among adults aged >=18 Years
* MHLTH: Mental health not good for >=14 days among adults aged >=18 Years
* PHLTH: Physical health not good for >=14 days among adults aged >=18 Years
* SLEEP: Sleeping less than 7 hours among adults aged >=18 Years


